Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-03-22 RMS/submission.tex
finished the proof

The following lemma is needed to prove the tick-invariance of our
defined propositions:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
Let $\cal R$ be a time-robust object-oriented specification with a
standard tick rule, let the infinity element \verb|INF| be the only
element in the time domain which is not a normal time value, and let
the time domain be \emph{linear}. Then $\cal R$ is tick-invariant with
respect to a set $P$ of atomic propositions if \verb|{|$t$\verb|}|
$\simeq_P$ \verb|{delta(|$t$\verb|,|$r$\verb|)}| for all $t,r$ with
$r<$ \verb|mte(|$t$\verb|)|.
\end{lemma}

@@

Condition~(ii) follows from the fact that $r+0=r$ and
$r$~\verb|monus|~$0=r$ for all $r$.
\end{IEEEproof}

>>>>

The following lemma is needed to prove the tick-invariance of our
defined propositions:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:tickinv}
Let $\cal R$ be a time-robust object-oriented specification with a
standard tick rule, let the infinity element \verb|INF| be the only
element in the time domain which is not a normal time value, and let
the time domain be \emph{linear}. Then $\cal R$ is tick-invariant with
respect to a set $P$ of atomic propositions if \verb|{|$t$\verb|}|
$\simeq_P$ \verb|{delta(|$t$\verb|,|$r$\verb|)}| for all $t,r$ with
$r<$ \verb|mte(|$t$\verb|)|.
\end{lemma}

@@

Condition~(ii) follows from the fact that $r+0=r$ and
$r$~\verb|monus|~$0=r$ for all $r$.

Condition~(iii). We must prove \deltaS{\deltaS{$s$}{$r$}}{$r'$} $=$
\deltaS{$s$}{$r+r'$} for all $r+r'\le$ \mteS{$s$}, with $s$ being a
system state of sort \verb|System|. Using the same symbols as in~(i),
we only consider the case $\mathit{ID}$\verb|::MaybeNat| in detail. 
By definition, the left side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{\deltaS{s}{r}}{r'}
\\  
= & \verb|(|\deltaTask{ID}{\deltaTask{ID}{L}{r}}{r'}
\\
  & \verb| |~T~STS~HW~\deltaIS{\deltaIS{ISRC}{r}}{r'}\verb|)|~\mbox{,}
\end{IEEEeqnarray*}
and the right side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{s}{r+r'}
\\  
= & \verb|(|\deltaTask{ID}{L}{r+r'}
\\
  & \verb| |~T~STS~HW~\deltaIS{ISRC}{r+r'}\verb|)|~\mbox{.}
\end{IEEEeqnarray*}
\deltaTask{$ID$}{\deltaTask{$ID$}{$L$}{$r$}}{$r'$} $=$
\deltaTask{$ID$}{$L$}{$r+r'$} holds by the associativity of $+$, while
\deltaIS{\deltaIS{$ISRC$}{$r$}}{$r'$} $=$ \deltaIS{$ISRC$}{$r+r'$}
holds since $(v$ \verb|monus| $r)$ \verb|monus| $r'$ $=$ $v$
\verb|monus| $(r+r')$ with $v$ of sort \verb|Time|. Thus
condition~(iii) holds.

Condition~(iv). We show that \verb|mte| of each instance of the
left-hand side of any instantaneous rule is $0$. For example,
considering the (conditional) rule \verb|interrupt-request|, with its
condition \verb|(|$ISRC$\verb|).timeout|, we know that the \verb|val|
of $ISRC$ equals $0$. Therefore,
\begin{IEEEeqnarray*}{Cl}
  & \mteS{L~T~STS~HW~ISRC}
\\  
= & \verb|minimum(|\mteTask{ID}{L}\verb|,|
\\
  & \verb|        |\mteIS{ISRC}\verb|,| \mteIr{HW}\verb|)|
\\
= & \verb|minimum(|\mteTask{ID}{L}\verb|,|~0~\verb|,| \mteIr{HW}\verb|)|
\\
= & 0~\mbox{.}
\end{IEEEeqnarray*}
The other rules can be similarly proved with their conditions. Hence
our model is time-robust by Lemma~\ref{l:timerobustness}.

Finally we prove the tick-invariance of the propositions used to
analyze our model, i.e. \verb|taskTimeout| and \verb|correct|. By
Lemma~\ref{l:tickinv}, we must prove
$\verb|{|s\verb|}|\simeq_P\verb|{|\deltaS{s}{r}\verb|}|$ with $s$ of
sort \verb|System| and $r<\mteS{s}$, that is, applying a tick rule
advancing $r$ time units will not change the value of each
proposition. Let $s$ be
$(\mathit{L}~\mathit{T}~\mathit{STS}~\mathit{HW}~\mathit{ISRC})$ and
$\verb|(|\mathit{HW}\verb|).getPc|=\mathit{ID}$.
\begin{itemize}
\item \verb|taskTimeout| holds if and only if $L$ contains an
  \verb|error|. Since \verb|delta| does not produce or eliminate
  \verb|error| in $L$, \verb|taskTimeout| holds in $s$ if and only if
  \verb|taskTimeout| holds in $\deltaS{s}{r}$ for any $r<\mteS{s}$,
  which means that our model is tick-invariant with respect to
  \verb|taskTimeout|.
\hide{ 
In our model, an \verb|error| cannot be eliminated by any operations,
hence, if \verb|taskTimeout| holds in $s$, it must hold in
$\deltaS{s}{r}$. On the other hand, \verb|error| can pop up only when
operation \verb|update_with_| is applied, i.e., only when the
instantaneous rule \verb|interrupt-handle| applies. If
\verb|taskTimeout| does not hold in $s$, it will not hold either in
$\deltaS{s}{r}$ for any $r<\mteS{s}$, since instantaneous rules cannot
apply in $\deltaS{s}{r'}$ for any $r'<\mteS{s}$ by time-robustness.}
\item The value of \verb|correct| depends on $ID$ and the status of
  each task in $L$. Similarly, \verb|delta| does not change $ID$ or
  the status of any task in $L$, hence \verb|correct| holds in $s$ if
  and only if \verb|correct| holds in $\deltaS{s}{r}$ for any
  $r<\mteS{s}$. It concludes the tick-invariance of \verb|correct|.
\end{itemize}

Therefore, by Theorem~\ref{t:completeness}, our approach using untimed
model checking to verify the schedulability and the correctness is
complete.
\end{IEEEproof}
